A non-additive generalization of probability theory\\for quantum mechanics and beyond

### 2025-07-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Cleanup first pages

\title{A non-additive generalization of probability theory \\for quantum mechanics and beyond}
\author{Gabriele Carcassi}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

@@

	%Given a generic space of ensembles, we can define the fraction capacity as the maximum fraction of a particular ensemble that can be understood as a mixture of ensembles from a given set. This gives a non-additive (i.e. fuzzy) measure that reduces to a probability (i.e. additive) measure for classical spaces and for quantum measurements. We can also define the state capacity as the exponential of the maximum entropy reachable by a mixture of ensembles from a given set. This is also a non-additive measure and reduces to the Liouville measure in classical mechanics and the Hilbert space dimensions for subspaces of quantum mechanics. Conceptually, it gives us a notion of probability that is both theory and interpretation independent. Mathematically, it gives us a measure theoretic generalization of probability. Physically, it may allow to find new way to understand current theory and tool to investigate new ones. The purpose of this paper is to show the core ideas and present questions that may be developed on the mathematical, physical and philosophical side.
\end{abstract}

@@

\section{TOC}

Overall goal: to create a generalization of probability theory and measure theory for general physical systems.

General question: how do we represent physical states? Different types of representations: talk about examples, ....

-- General problem

For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?

Separateness and orthogonality

Statistical quantities

Show that they reduce to standard things in classical mechanics and quantum mechanics.

-- Connection to Choquet theory

Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique

Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.

-- Connection to signed probability

TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})

Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).

Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex

Signed probability is unique, but does not characterize the convex set (ensemble space). Note: the other measures are defined on the extreme points. Knowing the extreme points means you know the convex space (i.e. take the hull). The non-additive measure does not characterize the extreme points, so it doesn't really characterize the space.

-- Our non-additive measure approach

State capacity

Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.

Fraction capacity as supremum of all possible measures

Additivity of fraction capacity (and state capacity).

Single decomposition and classical spaces and measurement contexts

Show that they reduce to standard things in classical mechanics and quantum mechanics.

-- Connection to fuzzy logic?

Literature search.

Possible connection to
* Choquet integral
* Sugeno integral

? Space of fuzzy measure like we have a space of additive measures?

>>>>

\title{A non-additive generalization of probability theory\\for quantum mechanics and beyond}
\author{Gabriele Carcassi}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

@@

	%Given a generic space of ensembles, we can define the fraction capacity as the maximum fraction of a particular ensemble that can be understood as a mixture of ensembles from a given set. This gives a non-additive (i.e. fuzzy) measure that reduces to a probability (i.e. additive) measure for classical spaces and for quantum measurements. We can also define the state capacity as the exponential of the maximum entropy reachable by a mixture of ensembles from a given set. This is also a non-additive measure and reduces to the Liouville measure in classical mechanics and the Hilbert space dimensions for subspaces of quantum mechanics. Conceptually, it gives us a notion of probability that is both theory and interpretation independent. Mathematically, it gives us a measure theoretic generalization of probability. Physically, it may allow to find new way to understand current theory and tool to investigate new ones. The purpose of this paper is to show the core ideas and present questions that may be developed on the mathematical, physical and philosophical side.

\end{abstract}

@@

\tableofcontents

%\section{TOC}
%
%Overall goal: to create a generalization of probability theory and measure theory for general physical systems.
%
%General question: how do we represent physical states? Different types of representations: talk about examples, ....
%
%-- General problem
%
%For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?
%
%Separateness and orthogonality
%
%Statistical quantities
%
%Show that they reduce to standard things in classical mechanics and quantum mechanics.
%
%-- Connection to Choquet theory 
%
%Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique
%
%Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.
%
%-- Connection to signed probability
%
%TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
%
%Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).
%
%Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex
%
%Signed probability is unique, but does not characterize the convex set (ensemble space). Note: the other measures are defined on the extreme points. Knowing the extreme points means you know the convex space (i.e. take the hull). The non-additive measure does not characterize the extreme points, so it doesn't really characterize the space.
%
%-- Our non-additive measure approach
%
%State capacity
%
%Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.
%
%Fraction capacity as supremum of all possible measures
%
%Additivity of fraction capacity (and state capacity).
%
%Single decomposition and classical spaces and measurement contexts
%
%Show that they reduce to standard things in classical mechanics and quantum mechanics.
%
%-- Connection to fuzzy logic?
%
%Literature search.
%
%Possible connection to
%* Choquet integral
%* Sugeno integral
%
%? Space of fuzzy measure like we have a space of additive measures?
